Efficient and Scalable Bipartite Matching with Fast Beta Linkage (fabl)

### 2022-09-11 ba_resubmit/kundinger_resubmit.tex
BK edits

Under the Fellegi Sunter model, we assume first, that comparison vectors are independent given their matching status, and second, that the match status of the record pairs are independent. This allows us to model the comparison data using mixture models in the following way:
\begin{align}
	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 1 \stackrel{iid}{\sim} \text{Multinomial}(\bm{m}), \\
	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \text{Multinomial}(\bm{u}), \notag\\
	& \Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(p), \notag
\end{align}
where $p$ is the proportion of records that match, and $\bm{m}$ is the set of parameters for the matching record pairs $m_{fl} = P(\gamma_{ij}^f = l|\Delta_{ij} = 1)$ for all fields $f$ and agreement levels $l$, and $\bm{u}$ is defined similarly for non-matching pairs. These $\bm{m}$ and $\bm{u}$ parameters are then used to calculate the likelihood ratio  
\begin{align}
	\label{eqn:wts}
	w_{ij} = \frac{P(\gamma_{ij} \mid \Delta_{ij} = 1)}{P(\gamma_{ij} \mid \Delta_{ij} = 0)}.
\end{align}
This ratio will be large when there is strong evidence of the pair being a match, and small otherwise.

@@

%The model can then be specified as follows:
With the \texttt{fabl} prior defined and our missingness assumptions addressed, we present the full model as follows: 
\begin{align*}
	\mathcal{L}(\bm{Z}, \Phi \mid \bm{\Gamma}) &= \prod_{j=1}^{n_1}  \prod_{i=1}^{n_2}\prod_{f=1}^{F}\prod_{l=1}^{L_f}\left[  m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}, \\
	\bm{m}_f &\sim \text{Dirichlet}(\alpha_{f0}, \ldots, \alpha_{f L_f}), \\
	\bm{u}_f &\sim \text{Dirichlet}(\beta_{f0}, \ldots, \beta_{f L_f}), \\
	Z_j | \pi &=
	\begin{cases} 
		\frac{1}{n_1}\pi,  & z_j \leq n_1; \\
		1-\pi, &  z_j  = n_1 + j; \\
	\end{cases} \\
	\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}),
\end{align*}
where we have used the condensed notation $\Phi = \{\bf{m}, \bf{u}\}$ to denote the set of parameters governing the probabilities of agreement levels for matching and non-matching record pairs.

>>>>

Under the Fellegi Sunter model, we assume first, that comparison vectors are independent given their matching status, and second, that the match status of the record pairs are independent. This allows us to model the comparison data using mixture models in the following way:
\begin{align}
	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 1 \stackrel{iid}{\sim} \mathcal{M}(\bm{m}), \\
	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \mathcal{U}(\bm{u}), \notag\\
	& \Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(p), \notag
\end{align}
where $\mathcal{M}$ and $\mathcal{U}$ are the distributions for matching and non-matching record pairs, $\bm{m}$ and $\bm{u}$ are their respective sets of parameters, and $p$ is the marginal probability that a record pair is a match. These $\bm{m}$ and $\bm{u}$ parameters are then used to calculate the likelihood ratio  
\begin{align}
	\label{eqn:wts}
	w_{ij} = \frac{P(\gamma_{ij} \mid \Delta_{ij} = 1)}{P(\gamma_{ij} \mid \Delta_{ij} = 0)}.
\end{align}
This ratio will be large when there is strong evidence of the pair being a match, and small otherwise.

@@

%The model can then be specified as follows:
With the \texttt{fabl} prior defined and our missingness assumptions addressed, we present the full model. Let $\bm{m} = (\bm{m}_1, \ldots, \bm{m}_F)$, where $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$ and $m_{fl} = P(\gamma_{ij}^f = l|Z_j = i)$ for all fields $f$ and agreement levels $l$. Let $\bm{u}$ parameters be defined analogously, with $u_{fl} = P(\gamma_{ij}^f = l|Z_j \neq i)$. Then, using the condensed notation $\Phi = \{\bf{m}, \bf{u}\}$, our model is:
\begin{align*}
	\mathcal{L}(\bm{Z}, \Phi \mid \bm{\Gamma}) &= \prod_{j=1}^{n_1}  \prod_{i=1}^{n_2}\prod_{f=1}^{F}\prod_{l=1}^{L_f}\left[  m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}, \\
	\bm{m}_f &\sim \text{Dirichlet}(\alpha_{f0}, \ldots, \alpha_{f L_f}), \\
	\bm{u}_f &\sim \text{Dirichlet}(\beta_{f0}, \ldots, \beta_{f L_f}), \\
	Z_j | \pi &=
	\begin{cases} 
		\frac{1}{n_1}\pi,  & z_j \leq n_1; \\
		1-\pi, &  z_j  = n_1 + j; \\
	\end{cases} \\
	\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}),
\end{align*}
%where we have used the condensed notation $\Phi = \{\bf{m}, \bf{u}\}$ to denote the set of parameters governing the probabilities of agreement levels for matching and non-matching record pairs.
